### 2023-02-20 Paper Two .qmd
Initial commit

---
title: "Paper Two"
author: 'SHAOHAN CHANG'
date: 02-20-2023
abstract: The investigation into how subway system openings affect urban air pollution, with a focus on particulate concentrations, is presented in this abstract. The study discovered that the opening of a subway system was associated with a 4% decrease in particulates in the area of the city center, which persisted over a four-year time horizon, in cities with higher initial pollution levels. In heavily polluted cities, this reduction was projected to have an annual external mortality benefit of about $1 billion. The findings imply that a significant portion of the cost of building subways may be offset by the reduced mortality caused by less air pollution. The results of this study can be repeated in other cities to learn more about the connection between air pollution and subway openings.
date-format: "MM/DD/YYYY"
format: pdf
bibliography: "bib.bib"
---

## Introduction

The paper investigates the impact of subway systems on air pollution in cities around the world. It uses two main data sources: a description of the world's subway systems and a measure of airborne particulates, Aerosol Optical Depth (AOD), recorded by satellites between February 2000 and December 2017. The analysis uses a comparison of changes in AOD within a city before and after the opening of a subway system to establish its causal effect.

The results show that the average effect of subway openings on AOD is a small decrease that cannot be distinguished from zero, but there is significant heterogeneity across cities. In the case of the 26 cities where AOD fell after the subway opened, the decrease was largest in cities whose initial level of AOD was above the median. The decrease in AOD levels was found to persist for at least 4 years. The results also indicate that subway ridership is a key factor in reducing AOD levels and that subway expansions beyond the initial line have small effects on AOD levels.

The study provides important information for policymakers considering the implementation of subway systems to mitigate air pollution. Based on the results, the authors estimate that a subway opening in an average city initially in the top half of the AOD distribution prevents 22.5 infant deaths and 500 total deaths per year, which is worth about \$43m and \$1b per year, respectively. The results suggest that subway systems may be cost-effective in reducing air pollution, particularly in cities with high initial levels of AOD. The study also sheds light on transportation behavior in developing countries, finding no evidence of differences between developing and developed world cities in their response to subways.

The study aims to investigate the effect of subways on urban air pollution by using data from a panel of cities. The subways data, used from Gonzalez-Navarro and Turner (2018) and updated to December 2017, define a subway as an electric-powered urban rail system isolated from interactions with vehicle and pedestrian traffic. The latitude, longitude, and date of opening of every subway station in the world were compiled manually between 2012 and 2014. The air pollution data used are based on remotely sensed measures of suspended particulates from Terra and Aqua satellites, providing daily measures of aerosol optical depth of the atmosphere at a 3 km spatial resolution from 2000 to 2017. The study considers the change in aerosol optical depth (AOD) in the period extending from 18 months before to 18 months after a subway opening, using a sample of 58 subway system openings between 2001 and 2016. The study also has ridership data for 42 of the 58 cities, with an average daily ridership of 130,000 people in the 18th month of operation. The study finds that ridership triples over the first three years of operation and begins to slow after three years. The average time construction began was 77 months prior to the opening.

The results of a study on the Aerial Optical Depth (AOD) within 10 km of city centers using satellite imagery from Terra and Aqua satellites. The study aimed to evaluate the relationship between the remotely-sensed AOD and ground-measured particulate matter (pm10 and pm2.5). The results show that AOD is a highly predictive measure of ground-measured particulate matter and the relationship is not sensitive to the exact region used to calculate city average AOD. The study used a monthly average of AOD readings within 10 km of the city center, calculated by averaging over all pixel-days of AOD readings that fall in this region during the month and weighting by the number of pixels observed each day. The results showed that in 2017, the average AOD reading within 10 km of a city center from the Aqua satellite was 0.40 and higher in Asian cities, whereas it was lower in European and North American cities. The second panel of Table 1 presents the AOD averages for 2000 using only the Terra satellite.

## Plot

```{r}
library(here)
library(dplyr)

subway_data <- read.csv(file = '/Users/lucas/Documents/paper2/Subway_Ridership_V4.csv')
head(subway_data)

subway_data<- subway_data %>%
  mutate(Ridership = ifelse(Month == ' ' , NA, Ridership))

```

```{r}
#| eval: true
#| warning: false
#| echo: false
#| label: fig-one
#| fig-cap: "Monthly Ridership in Santo Domingo, Dominican Republic"

library(dplyr)
library(ggplot2)
subway_data_monthly <- subway_data %>% 
  filter(ReferencePeriod != "Yearly")
ggplot(subway_data_monthly, aes(x = Month, y = as.numeric(Ridership)/1000, group = 1)) +
  coord_flip() +
  geom_point(color = "red") +
  labs(title = "Monthly Ridership in Santo Domingo, Dominican Republic",
       x = "Month", y = "Ridership",
       subtitle = "February 2000 - February 2020") +
  theme_minimal() +
  theme(plot.caption = element_text(size = 10))

```

```{r}
library(dplyr)
library(ggplot2)

# Filter the data to the desired year range and remove rows with missing values
year_range <- subway_data %>% 
  filter(Year >= 2000, Year <= 2010)

year_range

ggplot(year_range, aes(x = Month, y = as.numeric(Ridership)/1000)) +
  geom_line(color = "blue") +
  labs(title = "Subway Ridership in Santo Domingo, Dominican Republic",
       x = "Month", y = "Ridership",
       subtitle = "February 2000 - December 2010") +
  theme_minimal() +
  theme(plot.caption = element_text(size = 10))

```

## Conclusion

The paper discussed here is a study to investigate the impact of subway systems on air pollution in cities around the world. The study used two main data sources: satellite-recorded descriptions of the world's subway systems between February 2000 and December 2017 and measurements of the aerosol optical depth (AOD) of particulate matter in the air. The analysis used a comparison of changes in AOD in cities before and after the opening of the subway system to establish its causality.

The results show that the average effect of subway opening on AOD is a small decrease that cannot be distinguished from zero, but there is significant heterogeneity among cities. Among the 26 cities whose AOD decreased after the opening of the subway, the cities with the initial level of AOD higher than the median had the largest decrease. The decline in AOD levels was found to persist for at least 4 years. The results also show that subway ridership is a key factor in reducing AOD levels, and that subway expansion beyond the initial line has little effect on AOD levels.

The findings of this study provide important information for policymakers considering implementing a subway system to mitigate air pollution. The authors estimate that opening subways in an average city originally located in the upper half of the AOD distribution could prevent 22.5 infant deaths and 500 total deaths per year, worth about \$43 million and \$1 billion per year, respectively. The results suggest that subway systems may be cost-effective in reducing air pollution, especially in cities with high initial levels of AOD.

The study also sheds light on traffic behavior in developing nations, concluding that there are no differences between how developed and developing cities react to subways. This is a significant finding because developing nations may have a greater need for affordable air pollution solutions.

The Terra and Aqua satellites' remote sensing measurements of suspended particulate matter served as the basis for the study's data. Using satellite imagery from the Terra and Aqua satellites, the study evaluated the correlation between remotely sensed AOD and ground-based measurements of particulate matter (pm10 and pm2.5) within 10 km of the city center. The findings demonstrate that AOD is a highly reliable predictor of ground-measured particulate matter and that there is no significant relationship with the precise area used to derive the city-average AOD. The study made use of monthly averages of AOD readings for pixel days that occurred within a 10-kilometer radius of the city center during the month, weighted by the number of pixels observed on each day.

The study's use of changes in AOD levels to gauge the effect of the subway system on air pollution has some limitations. Although AOD is a trustworthy indicator of air pollution, it is only one facet of the intricate issue of urban air pollution. Future studies should look into how the subway system affects other pollutants and the overall quality of the air.

In conclusion, the studies covered in this article offer crucial details on how subway systems affect air pollution in urban areas around the globe. The findings imply that subway systems might reduce air pollution in cities with high initial levels of AOD at a reasonable cost. The findings of this study could be used by policymakers to decide whether to implement subway systems as a way to reduce air pollution. \## Reference

## Reference